Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
subseteq

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \di X \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

\[
	\di \sx = \di X \oplus \left(\bigoplus_{i
	= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subset \BQ \otimes \di X
\]
\end{defn}

@@

Our inductive arguments in Section ~\ref{sec:induction} will
usually have an inclusion $R \supset R'$ of log spin canonical
rings such that $R'$ is generated by elements $x_{d_i, i}$ and $R$
is generated over $R'$ by elements $y_j$. In these cases, it is
natural to consider term orders which treat these sets of variables
separately. We now present the \textbf{block term order} introduced
by Voight and Zureick-Brown \cite[Section 8.1]{vzb:stacky}.

@@

\todo{write below part}
For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial(I)$ to be the ideal generated by $\{\initial(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subset I$ such that 
+\[ \langle \initial(f_1),\dots,\initial(f_n) \rangle = \initial(I). \]

@@

\begin{lem}
\label{lem:admissible_subset}
If $(\sx',\Delta,L')$ is a log spin curve, $(\sx',L', J)$ is admissible and $W \subset J$ is any subset,
then $(\sx', L', W)$ is also admissible.
\end{lem}

@@

\noindent
and so we obtain an inclusion
\begin{align*}
	H^0(\sx', e_iL'-E_i + Q_i) \rightarrow H^0(\sx, e_iL - E_i) \subset 
H^0(\sx, e_iL).
\end{align*}

@@

Since $(*_0)$ holds by assumption, by induction on $i$, it suffices to show that if $(*_i)$ holds then so does $(*_{i+1}).$ 
Indeed, by applying Lemma ~\ref{lem:admissible_subset}, it follows $(\sx_i, \halfcan_i, J_{i+1})$ is admissible, as $J_{i+1} \subset J_i$. Then, applying Lemma ~\ref{lem:raise-stacky-order} $\frac{e_{i+1}-e_i}{2},$ times, for the fixed set $J_{i+1},$ we see $(*_{i+1})$ holds.
\end{proof}

@@

\begin{lem}
\label{lem:semicanonical_generation}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K + \Delta$. Let $s_1,s_2 \in H^0(X,2L)$ be two independent sections such that the vector subspace $V = ks_1 \oplus k s_2 \subset H^0(K)$ is basepoint free. Then, the map
\begin{align*}
	V \otimes H^0(nL) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 5$ and if {\rm(}$n = 4$ and $\Delta > 0${\rm)}.
In particular, $R_L$ is generated in degree at most $6$ if $\Delta = 0$ and in degree at most $5$ if $\Delta > 0$.
\end{lem}

@@

\begin{proof}[Proof of Lemma ~\ref{lem:semicanonical_generation}]
To show there are no new generators in degree at least $7$ if $\Delta = 0$ and in degree at least $6$ if $\Delta > 0$, it suffices to show that if $n \geq 5$ or if ($n = 4$ and $\Delta > 0$), the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, since $V= ks_1 \oplus k s_2 \subset H^0(K)$ is basepoint free, by the basepoint-free pencil trick, (see \cite[Lemma 2.6]{saint-donat:proj} for a proof,) we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

@@

\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ be two elements so that $k\phi(s_1) \oplus k\phi(s_2) = V \subset H^0(X,2L)$ is basepoint free.
For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
% Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ generate a base point free pencil $V = k \phi(s_1) \oplus k \phi(s_2)\subset H^0(X,2L)$.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since the restriction on $\deg f$ guarantees
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0,\\
	4 &\text{ if }\Delta > 0.
\end{cases}$$
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Choosing $g_i,h_i$ for $1 \leq i \leq n$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i \equiv s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

\begin{prop}
\label{prop:relation_11}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with 
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0.
\end{cases}$$
To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$ where $\phi(s_1),\phi(s_2) \in H^0(X,2L)$ are two sections so that $k\phi(s_1) \oplus k \phi(s_2) = V \subset H^0(K)$ is basepoint free. Consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, by the explicit isomorphism given in the proof of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Hence,
\begin{align*}
	s_1g + s_2h \equiv s_1(s_2\rho) + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}

>>>>

\todo{combine below two paragraphs with the corresponding
near-duplicate part from the background}
If $X$ is a curve, \todo{why do you start with a curve and then
move to a stacky curve? shouldn't it be the opposite order?} then
we say $\sx$ is a {\bf stacky curve} with {\bf coarse space} $X$
and {\bf stacky points} (also called {\bf{fractional points}}) $P_1,
\ldots, P_r$ with {\bf stabilizer orders} $e_1,\ldots, e_r \in \BZ_{
\geq 2}$ if 
\[
	\di \sx = \di X \oplus \left(\bigoplus_{i = 1}^r \left\langle 
	\frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X.
\]
\todo{Peter: Would it be better to say $Div(\sx)=(Div(X))[\frac{1}{e_i}P_i]_{1\le i\le r}$ or something like that? Aaron: I agree what is written isn't actually correct, maybe we should say the free direct sum over points not in the $P_i$, together with the fractional points. It's not a localization because we don't want to invert $P_i$.}
As is done in Voight and Zureick-Brown \cite{vzb:stacky} stacky curves may be formalized as stacks, but for the purposes of this paper, it suffices to view $\sx$ as a scheme whose associated divisor group allows a finite number of stacky points. \todo{I don't know if this is a good statement to make.  We aren't really viewing $\sx$ as a scheme.  Rather we're are determining information about its canonical rings through knowledge of scheme data.}  A {\bf log spin curve} is a triple $(\sx, \Delta, \halfcan)$ where $\Delta \in \di X$ is an effective divisor and $2\halfcan = K_X + \Delta + \sum_{i=1}^{r}\subhalf{e_i}P_i.$
The central object of study in this paper is the {\bf log spin
canonical ring} of $(\sx, \Delta, \halfcan)$ is
\begin{align*}
	R(\sx, \Delta, \halfcan) = \bigoplus_{d \geq 0} H^0(\sx,\lfloor d \halfcan \rfloor)
\end{align*}

@@

\[
	\di \sx = \di X \oplus \left(\bigoplus_{i
	= 1}^r \left\langle \frac{1}{e_i}P_i \right\rangle \right) \subseteq \BQ \otimes \di X
\]
\end{defn}

@@

Our inductive arguments in Section ~\ref{sec:induction} will
usually have an inclusion $R \supseteq R'$ of log spin canonical
rings such that $R'$ is generated by elements $x_{d_i, i}$ and $R$
is generated over $R'$ by elements $y_j$. In these cases, it is
natural to consider term orders which treat these sets of variables
separately. We now present the \textbf{block term order} introduced
by Voight and Zureick-Brown \cite[Section 8.1]{vzb:stacky}.

@@

\todo{write below part}
For a homogeneous polynomial $f \in k[x]_{\vec{a}}$, we define the \textbf{initial term} $\initial(f)$ to be the monomial in the support of $f$ that is largest with respect to $\prec$, and we define $\initial(0)=0$.  Let $I \subseteq k[x]_{\vec{a}}$ be a homogeneous ideal.  We define the \textbf{initial ideal} $\initial(I)$ to be the ideal generated by $\{\initial(f) : f \in I\}$.  A \textbf{Gr\"obner basis} (also called a \textbf{standard basis}) for $I$ is a set $\{f_1,\dots,f_n\} \subseteq I$ such that 
+\[ \langle \initial(f_1),\dots,\initial(f_n) \rangle = \initial(I). \]

@@

\begin{lem}
\label{lem:admissible_subset}
If $(\sx',\Delta,L')$ is a log spin curve, $(\sx',L', J)$ is admissible and $W \subseteq J$ is any subset,
then $(\sx', L', W)$ is also admissible.
\end{lem}

@@

\noindent
and so we obtain an inclusion
\begin{align*}
	H^0(\sx', e_iL'-E_i + Q_i) \rightarrow H^0(\sx, e_iL - E_i) \subseteq 
H^0(\sx, e_iL).
\end{align*}

@@

Since $(*_0)$ holds by assumption, by induction on $i$, it suffices to show that if $(*_i)$ holds then so does $(*_{i+1}).$ 
Indeed, by applying Lemma ~\ref{lem:admissible_subset}, it follows $(\sx_i, \halfcan_i, J_{i+1})$ is admissible, as $J_{i+1} \subseteq J_i$. Then, applying Lemma ~\ref{lem:raise-stacky-order} $\frac{e_{i+1}-e_i}{2},$ times, for the fixed set $J_{i+1},$ we see $(*_{i+1})$ holds.
\end{proof}

@@

\begin{lem}
\label{lem:semicanonical_generation}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2 L \sim K + \Delta$. Let $s_1,s_2 \in H^0(X,2L)$ be two independent sections such that the vector subspace $V = ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free. Then, the map
\begin{align*}
	V \otimes H^0(nL) \rightarrow H^0((n+2)L)
\end{align*}
is surjective if $n \geq 5$ and if {\rm(}$n = 4$ and $\Delta > 0${\rm)}.
In particular, $R_L$ is generated in degree at most $6$ if $\Delta = 0$ and in degree at most $5$ if $\Delta > 0$.
\end{lem}

@@

\begin{proof}[Proof of Lemma ~\ref{lem:semicanonical_generation}]
To show there are no new generators in degree at least $7$ if $\Delta = 0$ and in degree at least $6$ if $\Delta > 0$, it suffices to show that if $n \geq 5$ or if ($n = 4$ and $\Delta > 0$), the map
\begin{align*}
	H^0(nL) \otimes H^0(2L) \rightarrow H^0((n+2)L)
\end{align*}
is surjective. Indeed, since $V= ks_1 \oplus k s_2 \subseteq H^0(K)$ is basepoint free, by the basepoint-free pencil trick, (see \cite[Lemma 2.6]{saint-donat:proj} for a proof,) we obtain an exact sequence
$$\begin{tikzcd}
0 \ar r & H^0((n-2)L) \ar r & V \otimes H^0(nL) \ar{r}{f} & H^0((n+2)L)
\end{tikzcd}$$

@@

\begin{lem}
\label{lem:reducing_degree}
Suppose $X$ is a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K+\Delta$. Choose generators $x_1, \ldots, x_n$ of $R_L$ so that we obtain a surjection $\phi:k[x_1, \ldots, x_n] \twoheadrightarrow R_L$ with kernel $I_L$. Let $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ be two elements so that $k\phi(s_1) \oplus k\phi(s_2) = V \subseteq H^0(X,2L)$ is basepoint free.
For any $f \in k[x_1, \ldots, x_n]$ such that
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0
\end{cases}$$
there exist $g,h \in k[x_1, \ldots, x_n]_{d-2}$ so that $s_1g + s_2h = f \bmod J_{L,d}.$
\end{lem}
\begin{proof}
% Choose $s_1,s_2 \in k[x_1, \ldots, x_n]_2$ so that $\phi(s_1),\phi(s_2) \in H^0(2L)$ generate a base point free pencil $V = k \phi(s_1) \oplus k \phi(s_2)\subseteq H^0(X,2L)$.
By Lemma ~\ref{lem:generation_5}, $\deg x_i \leq 5$ for $1 \leq i \leq n$. Then, we may write $f = \sum_{i = 1}^{n}a_i x_i$ with $a_i \in k[x_1, \ldots, x_n]_{d-\deg x_i}$.
We will next show that for all $1 \leq i \leq n$ there exist $g_i, h_i \in k[x_1, \ldots, x_n]_{d - \deg x_i - 2}$ so that $a_i = s_1g_i + s_2h_i \bmod I_{L,\deg a_i}.$ 
By Lemma ~\ref{lem:semicanonical_generation},
\begin{align*}
	V \otimes H^0((\deg f-\deg x_i -2)L) \rightarrow H^0((\deg f-\deg x_i)L)
\end{align*}
is surjective, since the restriction on $\deg f$ guarantees
$$
\deg f- \deg a_i -2 \geq \begin{cases}
	5 &\text{ if }\Delta = 0,\\
	4 &\text{ if }\Delta > 0.
\end{cases}$$
In particular, we can write $\phi(a_i) = \phi(s_1) \alpha + \phi(s_1) \cdot \beta$. Choosing $g_i,h_i$ for $1 \leq i \leq n$ so that $\phi(g_i) = \alpha,\phi(h_i) = \beta$, we have $a_i \equiv s_1 g_i + s_2 h_i \bmod I_{L,\deg a_i}$, as claimed.

@@

\begin{prop}
\label{prop:relation_11}
Let $X$ be a curve of genus $g \geq 2$ and let $L$ be a divisor with $2L \sim K + \Delta$. Then $I_L$ is generated in degree at most $11$ if $\Delta = 0$ and degree at most $10$ if $\Delta > 0$.
\end{prop}
\begin{proof}
Suppose $f \in I_L$ with 
$$
\deg f \geq \begin{cases}
	12 &\text{ if }\Delta = 0\\
	11 &\text{ if }\Delta > 0.
\end{cases}$$
To complete the proof, it suffices to show $f \in I_{L,d}'$. By Lemma ~\ref{lem:reducing_degree}, it suffices to show $s_1g+s_2h \in I_{L,d}'$ where $\phi(s_1),\phi(s_2) \in H^0(X,2L)$ are two sections so that $k\phi(s_1) \oplus k \phi(s_2) = V \subseteq H^0(K)$ is basepoint free. Consider the map
$$\begin{tikzcd}
V \otimes H^0((\deg f - 2)L) \ar {r}{f} & H^0((\deg f)L),
\end{tikzcd}$$
we know that $\phi(s_1)\phi(g) + \phi(s_2) \phi(h) \mapsto 0.$
So, by the explicit isomorphism given in the proof of the basepoint-free pencil trick, as shown in the proof of \cite{saint-donat:proj}[Lemma 2.6], there exists some $\rho \in k[x_1, \ldots, x_n]$ so that $\phi(\rho) \in H^0((\deg f - 4)L)$ satisfies $\phi(g) = \phi(s_2)\phi(\rho)$ and $\phi(h) = -\phi(s_1)\phi(\rho).$ Therefore, $g \equiv s_2 \rho \bmod I_{k,d-2}$ and $h \equiv -s_1 \rho \bmod I_{k,d-2}$. Hence,
\begin{align*}
	s_1g + s_2h \equiv s_1(s_2\rho) + s_2(-s_1 \rho) \equiv 0 \bmod J_{L,d}.
\end{align*}
\end{proof}
